Aggressive Scheduling for Numerical Programs

### 2012-05-01 report/draft.tex
added a section. will return to this after I finish other commitment.

\begin{abstract}
   Certain classes of algorithms are still hard to optimize by modern compilers
   due to the differences between architecture features.
   Numerical algorithms, for example, perform better when taking advantage of features
   such as software pipelining and loop unrolling. Expert programmers are usually able to
   get the last drop of performance from the hardware, however, compilers cannot do the same
   for loss of generality. The SPIRAL system bridges this gap by generating machine specific code
   from high level algorithmic descriptions. We used the SPIRAL system to implement software
   pipelining for numerical algorithms with XXX (talk about scheduling). Our results show that
   ....
\end{abstract}

@@

The need to duplicate the loop code into three phases is a notorious disadvantage of software
pipelining, because it may increase the program's code size. Moreover, since software pipelining
is usually used in conjunction with loop unrolling, this enlarges the code even more.

\begin{verbatim}
+( ) Go into the background of software pipelining. How is it done?
+( ) How do we do it in SPIRAL?
     -( ) What parameters can we tune? (how much to delay an op)
     -( ) Formally what does this look like? What is our algorithm?
\end{verbatim}

\subsection{Enforcing a Schedule}
\begin{verbatim}
+( ) How do we enforce a schedule at the C level?
+( ) What does the code look like?
+( ) What is some example assembly code?

+( ) Would this hack always work? Either way it would be ideal to have a flag
     or pragma to allow the user to choose a schedule.

>>>>

\begin{abstract}
   Certain classes of algorithms are still hard to optimize by modern compilers
   due to the differences between architecture features.
   Numerical algorithms, for example, perform better when taking advantage of features
   such as software pipelining and loop unrolling. Expert programmers are usually able to
   get the last drop of performance from the hardware, however, compilers cannot do the same
   for loss of generality. The SPIRAL system bridges this gap by generating machine specific code
   from high level algorithmic descriptions. We used the SPIRAL system to implement software
   pipelining for numerical algorithms with XXX (talk about scheduling). Our results show that
   by unrolling and applying software pipelining at the C code level, we can
   improve performance by up to 30\% over unrolling alone.
\end{abstract}

@@

The need to duplicate the loop code into three phases is a notorious disadvantage of software
pipelining, because it may increase the program's code size. Moreover, since software pipelining
is usually used in conjunction with loop unrolling, this enlarges the code
even more. While this can be an issue on processors with little to no
instruction cache, in practice on most modern systems this is not an issue.

\begin{verbatim}
+( ) Go into the background of software pipelining. How is it done?
+( ) How do we do it in SPIRAL?
\end{verbatim}

\subsection{Enforcing a Schedule}
\begin{verbatim}
+(X) How do we enforce a schedule at the C level?
+( ) What does the code look like?
+(X) Would this hack always work? Either way it would be ideal to have a flag
     or pragma to allow the user to choose a schedule.

\end{verbatim}
Most modern compilers do not preserve instruction ordering imposed at the.
input language level, and instead rely on general purpose methods such
as \emph{list scheduling} to order instructions. In the common case this frees
the developer from the task of determining an efficient schedule, but in our
case this feature serves as a roadblock. While the GCC~\cite{} provides flags for
disabling instruction scheduling it comes at the cost of overall
performance because we only plan to schedule loop code. Moreover, our plan
involves using the Intel Compiler~\cite{}, thus we need a general solution for
scheduling at the C code level.

In order to enforce a schedule at the C code level we need to prevent the
compiler from imposing its own scheduling. A compiler must preserve the
semantics of the input code, so to take advantage of this fact we place a
label before each instruction with a \emph{goto} nested in
a \emph{switch-case} block. We declare the value for the \emph{switch}
statement as a global variable set such that the code will always jump to the
first instruction. Initially, we simply embedded each instruction in
a \emph{switch-case} block, but both the GCC and Intel compiler would create
multiple code paths and schedule each path individually. Figure~\ref{} shows
an example of our method in practice.

For our purposes this technique proved itself effective for enforcing a
schedule, but in some cases the GCC would add superfluous instructions. Thus
in the future a more general approach, such as a compiler directive, would be
advantageous.
